A \emph{Grimm} idea

### 2016-04-30 paper/main.tex
finished first version of simulation proof

In order to describe the movement of an agent we introduce the notion of an
exploration sequence as $e_{0}, e_{1},\dots$ which models relative movements
as follows. We use a sequence of variables $l_{0},l_{1},\dots$ which denotes
the port a vertex was entered through. And the next vertex is determined as the
one connected to the port $(l_{i} + e_{i}) \mod d_{v_{i}}$. We define
$l_{0} = 0$ and the notion $(e_{0},\dots,e_{k})^{\ast}$ for periodic sequences.
It can be shown that a periodic sequence produces a closed walk through a
graph.
\begin{proof}
  We argue that there is only a finite amount of triples of the
  form $(v,l,i)$ where currently the agent is on vertex $v$, entered
  this vertex through port $l$ and is currently at index $i$ of the exploration
  sequence. For two triples $(v',l',i')$ and $(v'',l'',i'')$ that both lead to
  $(v,l,i)$ it is easy to see, that $(v',l',i') = (v'',l'',i'')$. Thus, for
  every triple there is a unique predecessor and successor which implies due to
  the finiteness of the triples $(v,l,i)$ and that the agent moves infinitely
  it follows that the first triple has to be valid infinitly often. This means
  a periodic exploration sequence yields a closed walk.
\end{proof}

@@

  This simulation yields at most $3\Delta n^{2}$ many nodes and can be emulated
  with $\mathcal{O}(1)$ many memory space since it is only needed to store
  the second component of the inserted nodes which is bounded by $3\Delta$.
  Furthermore $A$ can be run on this constructed Graph and only needs
  $\mathcal{0}(\log n)$ space since $\mathcal{O}(\log(3\Delta n^{2})) =
  \mathcal{O}(\log(3) + \log(\Delta) + 2\log(n)) = \mathcal{O}(\log n)$.
  Which yields the required $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble
  machine.
\end{proof}

In order to obtain the desired boundary of $\log\log n$ space and $\log\log n$
pebbles we show that a pebble machine can be simulated by storing its tape
content by pebbles that are placed on a closed walk.
\begin{thm}
  For every graph $G$ a $(s,p,2m)$ pebble machine $M$ can be simulated by
  a $(cs,p + c,m)$ pebble machine that simulates the walk of $M$ or explores
  $G$.
\end{thm}
\begin{proof}
  The main idea of this simulation is to store the tape of $M$ by pebbles
  placed on a closed walk in the graph. We show that every node on this closed
  walk can be given an index and a pebble placed on a node represents the
  binary representation of the index as the content of $M$'s tape. Again we
  refer to the following picture for illustration: \todo{add picture}
  We will first introduce the algorithms used to implement the simulation and
  afterwards we will argue why this simulation respects the given space
  boundarys.  Let $M_{\text{Walk}}$ be a pebble machine that yields a closed
  walk on $G$. (Such a pebble machine exists due to Theorem
  \ref{thm:pebblewalk}.) In order to manage the simulation we use some

@@

  $P_{M} = \left\{p_{1},p_{2},\dots\right\}$.

@@

      there are no more vertices on the walk (we will see that this can be
      computed since the used walk is of constant size) it simply executes

@@

      the position of the bit that should be read. The walk of
      $M_{\text{walk}}$ is only considered to a certain length (as we will see
      later) which also determines the blocks in which $M$'s tape is separated.
      (The index of a vertex is used to encode one block.) Therefore we can use
      this length of the walk which pebble encodes the corresponding block
      which is obtained.  And then the actually needed bit of the index is
      returned.

@@

      $\Thead$'s bit is obtained (by \emph{ReadBit()}). And if it is actually
      changed the corresponding pebble is moved to store the new value of the
      memory cell. Thus, if the bit is flipped from 0 to 1 we increase $p_{i}$
      where $i$ is the encoded memory cell index by $2^{\Thead\bmod m_{1}}$
      where $m_{1}$ encodes the length of the considered walk. And if the bit
      is flipped from 1 to 0 we decrease $p_{i}$'s index by
      $2^{\Thead\bmod m_{1}}$.

@@

  These are the procedures used to simulate $M$. We will now discuss the space
  boundarys and afterwards conclude with showing how the simulation actually
  operates. Recall that $M$ is a $(s,p,2m)$ pebble machine. For the simulating

@@

>>>>

In order to describe the movement of an agent we introduce the notion of an
exploration sequence as $e_{0}, e_{1},\dots$ which models relative movements
as follows. We use a sequence of variables $l_{0},l_{1},\dots$ which denotes
the port a vertex was entered through. And the next vertex is determined as the
one connected to the port $(l_{i} + e_{i}) \mod d_{v_{i}}$. We define
$l_{0} = 0$ and the notion $(e_{0},\dots,e_{k})^{\ast}$ for periodic sequences.
It can be shown that a periodic sequence produces a closed walk through a
graph.
\begin{proof}
  We argue that there is only a finite amount of triples of the
  form $(v,l,i)$ where currently the agent is on vertex $v$, entered
  this vertex through port $l$ and is currently at index $i$ of the exploration
  sequence. For two triples $(v',l',i')$ and $(v'',l'',i'')$ that both lead to
  $(v,l,i)$ it is easy to see, that $(v',l',i') = (v'',l'',i'')$. Thus, for
  every triple there is a unique predecessor and successor which implies due to
  the finiteness of the triples $(v,l,i)$ and because the agent moves
  infinitely that the first triple has to be valid infinitly often.
  This means a periodic exploration sequence yields a closed walk.
\end{proof}

@@

  This simulation yields at most $3\Delta n^{2}$ many nodes and can be emulated
  with $\mathcal{O}(1)$ many memory space since it is only needed to store
  the second component of the inserted nodes which is bounded by $3\Delta$.
  Furthermore $A$ can be run on this constructed Graph and only needs
  $\mathcal{0}(\log n)$ space since $\mathcal{O}(\log(3\Delta n^{2})) =
  \mathcal{O}(\log(3) + \log(\Delta) + 2\log(n)) = \mathcal{O}(\log n)$.
  Which yields the desired $(\mathcal{O}(1),0,\mathcal{O}(\log n))$ pebble
  machine.
\end{proof}

In order to obtain the desired boundary of $\log\log n$ space and $\log\log n$
pebbles we show that a pebble machine can be simulated by storing its tape
content by pebbles that are placed on a closed walk.
\begin{thm}
  \label{thm:simulation}
  For every graph $G$ a $(s,p,2m)$ pebble machine $M$ can be simulated by
  a $(cs,p + c,m)$ pebble machine that simulates the walk of $M$ or explores
  $G$.
\end{thm}
\begin{proof}
  In this proof we will focus on the simulation and we leave out a thorough
  analysis of the space requirements of the simulation since it is rather
  technical but does not add a lot to the idea. Nevertheless we encourage
  interested readers to see \cite{pebbles} for the details.

  The main idea of this simulation is to store the tape of $M$ by pebbles
  placed on a closed walk in the graph. We show that every node on this closed
  walk can be given an index and a pebble placed on a node represents the
  binary representation of the index as the content of $M$'s tape. Therefore we
  first off introduce a variable $m_{1}$ which is dependent on the simulated
  pebble machine $M$ (further requirements on $m_{1}$ and how to obtain it
  can be seen in \cite{pebbles}). We will now take the tape of $M$ and separate
  it into $2m/m_{1}$ blocks of the size $m_{1}$. And each of these blocks is
  encoded by the placement of a pebble on a closed walk $\omega$. So we use
  pebble $p_{0}$ to encode the first $m_{1}$ bits, the pebble $p_{1}$ used to
  encode the next $m_{1}$ bits and so forth. Thus, the pebble $p_{i}$ encodes
  the bits $im_{1},\dots,(i+1)m_{1}-1$. It is therefore necessary that every
  pebble can encode $2^{m_1}$ different states. Therefore we use a closed walk
  $\omega$ which visits at least $2^{m_1}$ distinct vertices yielding $2^{m_1}$
  many \enquote{states} for each pebble (by its placement on this closed walk).
  Therefore we can use the binary representation of the indices of the nodes
  on which the pebbles are placed as the content of the encoded memory block.
  The following picture gives an illustration of this.\todo{add picture}

  In the following we show how this simulation can be managed.
  Let $M_{\text{Walk}}$ be pebble machine that yields $\omega$ on $G$.
  $M_{\text{Walk}}$ can be implemented as $(\mathcal{O}(1),0,\mathcal{m_1})$
  pebble machine by Theorem \ref{thm:pebblewalk}.
  In order to manage the simulation we use some

@@

  $P_{M} = \left\{p_{0},p_{1},\dots,p_{2m/m_{1}-1}\right\}$.

@@

      there are no more vertices on the walk (which means that already
      $2^{m_1}$ distinct vertices have been traversed; the number of distinct
      vertices on the walk is given by $\Tid$) it simply executes

@@

      the position of the bit that should be read. Therefore the pebble
      $p_{\lfloor \Thead/m_{1}\rfloor}$ encodes this bit. The \enquote{state}
      of this pebble can be obtained with
      \emph{GetPebbleId($p_{\lfloor \Thead/m_{1}\rfloor}$)} and from the
      encoded value the $(\Thead\mod m_{1})$-th bit is the one under $M$'s
      head.

@@

      $\Thead$'s bit is obtained (by \emph{ReadBit()}). Additionally we
      identify the encoding pebble $p_{i}$ as already seen for
      \emph{ReadBit()} and let be $j = \Thead\mod m_{1}$ the offset of the bit
      within the memory block. If the bit actually flips the value of the
      memory block changes by $2^j$ (for the flip from 1 to 0 the value
      decreases and for the flip from 0 to 1 it increases). Therefore $p_{i}$
      is put either $2^j$ indices up or down which can be easily done by using
      \emph{PutPebbleAtId(\dots)}.

@@

  These are the procedures used to simulate $M$'s tape. Furthermore we keep
  $M$'s states in the state space of the simulation. To conclude the simulation
  we have to consider what happens if $M$ moves from the current vertex to
  another vertex $v'$. Then the pebble $p_{tmp}$ is dropped on $v'$ and the
  pebbles are moved to the same index they have in $\omega$ for a new closed
  walk $\omega'$ (obtained by $M_{\text{Walk}}$ simulated starting in $v'$;
  this walk is also ensured to have at least $2^{m_1}$ distinct vertices).

@@

As Theorem \ref{thm:pebblewalk} yields a $(c,0,c\cdot\log n)$ pebble machine
that explores a given graph we can now repeatedly apply the simulation
from Theorem \ref{thm:simulation} (precisely $\log\log n$ times) and end up
with a $(c'^{\log\log n}\cdot c, \log\log n\cdot c',
\frac{c}{2^{\log\log n}}\log n) =
(c'^{\log\log n}\cdot c, \log\log n\cdot c', c)$ pebble machine exploring the
graph.
